---
title: "R Notebook"
author: "Mostafa"
date: 3/28/2021
output: html_notebook
---

---

```{r}
library (TSA) 
library(tseries)
```

# 1.
```{r}
# 1.
set.seed(5)
AR1 <- arima.sim(list(order = c(1,0,0), ar = -0.6), n = 1000)  # phi = -0.6
MA1 <- arima.sim(list(order = c(0,0,1), ma = -0.8), n = 1000)  # theta = 0.8

par(mfrow=c(2,2))

acf (AR1, main="Sample ACF,  AR(1), phi = -0.6")
pacf(AR1, main="Sample PACF, AR(1), phi = -0.6")

acf (MA1, main="Sample ACF,  MA(1), theta = 0.8")
pacf(MA1, main="Sample PACF, MA(1), theta = 0.8")

eacf(AR1) #, main="Sample PACF, AR(1), phi = -0.6"
eacf(MA1) #, main="Sample PACF, MA(1), theta = 0.8"
```
**Answer**: 
Yes. The results based on ACF, PACF and EACF consistent with the true model for the time series. The reasons for each of them are as below step by step.        
- Sample ACF  MA(1): A spike at lag 1 ==> A moving average of order 1 (q=1)                         
- Sample PACF AR(1): A spike at lag 1 ==> An Autoregressive process of order 1 (P=1)              
- Sample EACF AR(1): A wedge of zeros with tip at (1,0) ==> ARMA(1,0)                 
- Sample EACF MA(1): A wedge of zeros with tip at (0,1) ==> ARMA(0,1)        

---      

# 2.

## 2(a)
```{r}
# 2(a)
gasprices <- ts(read.table("gasprices.txt", quote="\"", comment.char=""))
par(mfrow=c(1,2))
plot(gasprices, ylab="Average prices (US dollars per gallon)",xlab="Time",type="o"
     , main="Average gas price, USA")
acf(gasprices,main="Sample ACF")

# ADF test
adf.test(gasprices)
```
**Answer**:    
- Time series plot ==> Mean is not zero and a linear upward trend ==> Non-stationary         
- Sample ACF ==> Ddecays slowly ==> Non-stationary         
- adf.test(gasprices): p-value(0.5492) ==> does not reject the null hypothesis. ==> There is insufficient evidence to conclude that the process is stationary.           

## 2(b)
```{r}
# 2(b)
par(mfrow=c(1,2))
plot(diff(gasprices),ylab="First differencing (gasprices)",xlab="Time",type="o"
     , main="diff() Average gas price, USA")
acf(diff(gasprices),main="Sample ACF: 1st differences(gasprices)")

# ADF test
adf.test(diff(gasprices))
```

**Answer**:    
- Time series plot of first differencing ==> It is better then the time series plot but it is not still mean zero, the points hang together and there are some big fluctuations. In my view it would not stationary.                     
- Sample ACF: It has a big spike at lag 1 and a small one at lag 15. As the spike at lag 15 is not an early lag, maybe it is a false positive because od an outlier. So, It is similar to an MA(1) process.                      
- adf.test(): p-value(0.01) is small ==> rejects the null hypothesis ==> There is sufficient evidence to formally conclude that the process is stationary.         

## 2(c)
```{r}
# 2(c)
par(mfrow=c(1,2))
pacf(diff(gasprices),main=" Sample PACF: 1st differences(gasprices)")
eacf(diff(gasprices))
```
**Answer**:    
- Sample PACF: A spike at lag 1 ==> AR(1) process              
- eacf(): a wedge of zeros with tip at (1,1) ==> The sample EACF largely agrees with the theory ARMA(1,1), which says that there should be a wedge of zeros with tip at (1,1). Now, we can make sure that the 2(b) was a MA(1) too.            

