---
title: "5160 - hw7"
author: "Mostafa"
date: 4/13/2021
output: 
    html_notebook:
        toc: yes
---

--- 

# 1.a. Transforming / Differencing       
```{r}
library(TSA) # Load the TSA package
library(tseries) # Load the tseries package
```

## TS Plot                
```{r}
# Load and plot the data
gasprices<- ts(read.table(file = "gasprices.txt"))
plot(gasprices, ylab = "Price (USD)", xlab = "Week", type = "o", main = "Average price for regular gasoline" )
```

## Transformation if nonconstant variance             
- No non-constant variance in TS plot ==> No need to transformation.
- CI lambda: .6-1.8 ==> No common transformation in this interval  

```{r}
# Profile log-likelihood of lambda (for Box Cox transformation)
# This can be computationally intense (time consuming)
BoxCox.ar(gasprices,method="burg")
```

## adf.test(): H0: non-stationsry     
The large p-value here (`0.5492`) does not reject the null hypothesis. There is insufficient evidence to conclude that the `gasprices` process is stationary.
```{r}
adf.test(gasprices)
```

## acf(): decay very slowly?    
**It seems that differencing is needed.**        
```{r}
acf(gasprices,main="Sample ACF")
```

## first differencing (d=1) and acf ==> MA(q=1)
```{r}
par(mfrow=c(1,2))
plot(diff(gasprices),ylab="First Difference of Gasoline Prices", xlab="Time",type="o")
acf(diff(gasprices),main="Sample ACF")
```

## adf.test() ==> stationary?      
H0: non-stationarity versus H1: stationarity    
`p-value = 0.01` ==> The null hypothesis is rejected. ==> Stationary 

```{r}
adf.test(diff(gasprices))
```
**Note**: 
What does "p-value smaller than printed p-value" mean?      
It means that this algorithm shows `p-value=0.01` if it is smaller or eual to `0.01`.

## pacf() ==> AR(p=1)
```{r}
pacf(diff(gasprices),main="Sample PACF")
```

## eacf() ==> ARMA(p,q)
```{r}
eacf(diff(gasprices))
```

## ARIMA(p,d,q) ==> ARIMA(1,1,1)    
- diff() ==> d=1
- acf(q=1)
- pacf(p=1)
- eacf(p=1,q=1)

**So, this model would be ARIMA(1,1,1)**

---    

# 1.b. Fitting a model

## Fit a model using maximum likelihood
```{r}
# Fit the model using ARIMA(1,1,1)
arima(gasprices,order=c(1,1,1),method='ML') # maximum likelihood
```

## To automatically check the significancy     
```{r}
a = arima(gasprices,order=c(1,1,1),method='ML') # maximum likelihood
# To automatically check the significancy
a$coef/sqrt(diag(a$var.coef))
```

## final model    
As MA(1) is not significant (between -2 and 2), I go for ARIMA(1,1,0). As we can see, this model is more significant.

```{r}
arima(gasprices,order=c(1,1,0),method='ML') # maximum likelihood
```

```{r}
a = arima(gasprices,order=c(1,1,0),method='ML') # maximum likelihood
# To automatically check the significancy
a$coef/sqrt(diag(a$var.coef))
```
**Answer**:      
The final model is $\nabla Y_{t} = 0.4878 \ Y_{t-1} + e_{t}$      
The white noise variance estimate is $\sigma^2 = 0.002344$.     

---  

# 1.c. The diagnostic checks
```{r}
gasprices.ma1.fit = arima(gasprices,order=c(1,1,0),method='ML') # maximum likelihood
gasprices.ma1.fit
```

## Normality check: qq plot        
- Residuals are distributed almost evenly around the the zero line.         
- Histogram is almost bell-shaped.      
- qqplot: The points are centered and most of the are on the reference line.     

**So, we can conclude that it is normaly distributed**.  
```{r}
par(mfrow=c(2,2))
plot(gasprices, ylab = "Price (USD)", xlab = "Week", type = "o", main = "Average price for regular gasoline" )
plot(rstandard(gasprices.ma1.fit),xlab="Time",ylab="Standardised residuals",type='o')
abline(h=0)
hist(rstandard(gasprices.ma1.fit),xlab="Standardised residuals",main="")
qqnorm(rstandard(gasprices.ma1.fit),main="")
qqline(rstandard(gasprices.ma1.fit))
```
## Normality check: Shapiro-Wilk tests    
H0 : the (standardized) residuals are normally distributed.         

- As `p-value = 0.006841`, the null hypothesis is rejected. And we can conclude that it is not normally distributed. However I prefer to ignore this result and trust qq plot. Because the tails are less important in a distribution. 
```{r}
# Shapiro-Wilk tests
shapiro.test(rstandard(gasprices.ma1.fit))
```

## Independency check: RUNS TEST
H0 : the (standardized) residuals are independent      
- A small number of runs would indicate that neighboring values are positively dependent and tend to
hang together over time.            
- Too many runs would indicate that the data oscillate back and forth across their median. This suggests
that neighboring residuals are negatively dependent.           
- Therefore, either too few or too many runs lead us to reject independence.           

**Conclusion**:     
- As the `$observed.runs = 71` is close to `$expected.runs = 73.41379`, we conclude that this amount of runs is not too many or small. So, we can trust the `$pvalue`.    
- As the `$pvalue = 0.749` is large, we can not reject the null hypothesis (residuals are independent). It means that there is not enough evidence to conclude that it is not independent.    

```{r}
# runs test
runs(rstandard(gasprices.ma1.fit))
```


## To check the quality of fit: acf() of the residuals                     
To further check the adequacy of a fitted model, we use acf() of the residuals. The lags are supposed to be approximately uncorrelated. It means that none of the sample autocorrelations fall outside the error bounds.     

**Conclusion**: As none of the sample autocorrelations fall outside the error bounds, the lags are approximately uncorrelated. Thereofre, the quality of this fit is acceptable.            
```{r}
acf(rstandard(gasprices.ma1.fit),main="Sample ACF for MA(1) residuals")
```

## To ckeck the Overfitting: ARIMA(p+1,d,q) & ARIMA(p,d,q+1)          

IMPORTANT: When overfitting an ARIMA(p, d, q) model, we consider the following two models:       
- (a) ARIMA(p+1,d,q)       
- (b) ARIMA(p,d,q+1)      

To ckeck:      
Main model: ARIMA(1,1,0)         
Check 1: ARIMA(2,1,0)        
Check 2: ARIMA(1,1,1)      
```{r}
arima(gasprices,order=c(2,1,0),method='ML') # maximum likelihood
arima(gasprices,order=c(1,1,1),method='ML') # maximum likelihood
```

**Analysis**:    

Always: distribution (approximately 1.96)      
$-0.0982 \ \pm \ 1.96 \ * 0.0840 \ = \ (-0.26284, 0.06644)$: This CI includes zero. Therefore, $\phi_{2}$ is not
statistically different than zero, which suggests that the AR(2) model is not necessary.       

 
$-0.1713 \ \pm \ 1.96 \ * 0.1479 \ = \ (-0.461184, 0.118584)$: This CI includes zero. Therefore, $\theta_{1}$ is not
statistically different than zero, which suggests that the MA(1) model is not necessary.


