---
title: "5160 - HW8"
author: "Mostafa"
date: 4/23/2021
output: 
    html_notebook:
          toc: yes
---

---

# 1.      
The `gasprices.txt` data set includes average price (US dollars per gallon) for regular gasoline in the United States; there are **145** weekly observations** collected from 1/5/2009 to 10/10/2011 (Source: RajonColes, Fall 2011). Use only the first n=**140** weekly observations to fit models and make predictions.

```{r cars}
library(TSA) 
library(tseries) 
```

## 1.a.      
- Calculate MMSE forecasts and prediction intervals for next 5 weeks (week 141 to week 145) using an ARIMA(0,1,1) model.     
- Display the forecasts and prediction intervals visually.

```{r}
gasprices <- ts(read.table("gasprices.txt", quote="\"", comment.char=""),start = 1, end = 140)

##  ARIMA(0,1,1) model
gasprices.arima011.fit = arima(gasprices,order=c(0,1,1),method='ML', n=140)
gasprices.arima011.fit

# Prediction and standard error
gasprices.arima011.fit.predict = predict(gasprices.arima011.fit, n.ahead=5)
round(gasprices.arima011.fit.predict$pred,3)

# Original series starts at week = 1
# Note: Argument n1=1 starts plot at 1
# Note: Argument pch=16 produces a small black circle (for MMSE forecasts)
plot(gasprices.arima011.fit, n.ahead=5, col='red', type='b', pch=16, n1=1, 
     ylab="Average price (USD/Gallon)", 
     xlab="weeks")

# show the prediction intervals visually
result.arima.011 = plot(gasprices.arima011.fit, n.ahead=5, col='red', type='b', pch=16, n1=1, 
                        ylab="Average price (USD/Gallon)", 
                        xlab="weeks", 
                        Plot = FALSE)
result.arima.011
```

## 1.b.    
- Calculate MMSE forecasts and prediction intervals for next 5 weeks (week 141 to week 145) using an ARIMA(1,2,0) model.        
- Display the forecasts and prediction intervals visually.

```{r}
##  ARIMA(1,2,0) model
gasprices.arima120.fit = arima(gasprices,order=c(1,2,0), method='ML', n=140)
gasprices.arima120.fit

# Prediction and standard error
gasprices.arima120.fit.predict = predict(gasprices.arima120.fit, n.ahead=5)
round(gasprices.arima120.fit.predict$pred, 3)

# Original series starts at week = 1
# Note: Argument n1=1 starts plot at 1
# Note: Argument pch=16 produces a small black circle (for MMSE forecasts)
plot(gasprices.arima120.fit, n.ahead=5, col='blue', type='b', pch=16, n1=1, 
     ylab="Average price (USD/Gallon)", 
     xlab="weeks")

# show the prediction intervals visually
result.arima.120 <- plot(gasprices.arima120.fit, n.ahead=5, col='blue', type='b', pch=16, n1=1, 
                         ylab="average price (US dollars per gallon)", 
                         xlab="weeks", 
                         Plot = FALSE)
result.arima.120
```
## 1.c.     
Since we already know the observed values from week 141 to week 145, we want to compare which model is better in prediction.     
For the above two models:     
- calculate    
- compare    
the mean squared forecast error (MSFE) using the following formula:     
$$MSFE = \frac {1}{5} \Sigma_{l=1}^{5} \{Y_{140+l} - \hat Y_{t}(l) \}^2$$
```{r}
gasprices.141.145 <- ts(read.table("gasprices.txt", quote="\"", comment.char=""),
                start = 141, 
                end = 145)

arima011 <- round(gasprices.arima011.fit.predict$pred, 3)
arima011
MSEF.arima011 = (1/5) * sum( (arima011 - gasprices.141.145)^2 )
MSEF.arima011

arima120 <- round(gasprices.arima120.fit.predict$pred, 3)
arima120
MSEF.arima120 = (1/5) * sum( (arima120 - gasprices.141.145)^2 )
MSEF.arima120 

if (MSEF.arima011 < MSEF.arima120) {
   print("ARIMA(0,1,1) is better than ARIMA(1,2,0)")
   }else{
   print("ARIMA(1,2,0) is better than ARIMA(0,1,1)")
   }
```

# Conclusion    

### ARIMA(0,1,1): MSFE = 0.0399142
### ARIMA(1,2,0): MSFE = 0.1119598
### 0.0399142 < 0.1119598 ==> ARIMA(0,1,1) is better that ARIMA(1,2,0)

