---
title: "5160_HW-5"
author: "Mostafa"
date: 3/16/2021
output: html_notebook
---

---

## 1     

### 1.a.
```{r}
library(TSA)
data(milk)
plot(milk, type = "o")
```
**Answer**:      
- Increasing trend: Yes          
- mean: increasing              
- variance: almost constant                
- Seasonality trend: Somehow               
- Stationary: No           

### 1.b. 
```{r}
diff_milk <- diff(milk)
plot(diff_milk, ylab="diff_milk", type = "o")
```
**Answer**:      
- Increasing trend: Not anymore             
- mean: almost zero               
- variance: almost constant             
- Seasonality trend: Somehow             
- Stationary: It seems still not                          

The increasing trend has been removed compared to the original data but still could be following a seasonal trend.
when compared to the inital gragh the second looks more close to stationary.   


### 1.c.

```{r}
fit1 <- lm(diff_milk~time(diff_milk))
fit1
```
**Answer**:    
The slope `-0.009603` is very low. So, we can conclude that it is almost zero mean.

```{r}
plot(fit1$residuals, ylab = "Residuals", type = "o")
abline(h = 0)
```
**Answer**:    
The points are evenly distributed around the abline=0. So, we can conclude that the variance is constant. However, there are some spikes that might be because of seasonality trend.So, it is hard to say that the residuals are independent. Finally, I think the residuals are not independent.

```{r}
qqnorm(rstudent(fit1))
qqline(rstudent(fit1))
```
**Answer**:    
Not only is not it normal, but also we can see some level of seasonality in it.


```{r}
shapiro.test(rstudent(fit1))
```
**Answer**:    
From pvalue we conclude that the standardized residuals are not normally distributed. 


```{r}
runs(rstudent(fit1))
```
**Answer**:    
first differences data: not normal & not independent

```{r}
qqnorm(diff(milk))
qqline(diff(milk))
       
plot(diff(milk),main="Differencing",type="o")
abline(h=0)
# Runs test to check the independency 
runs(diff(milk))
acf(diff(milk),main="Differencing")
```
**Answer**:    
There are some spikes fall out of the margin errors cut off. So, these likely does not resemble a white noise process.

---

## 2    
```{r}
set.seed(1)
AR1 <- arima.sim(list(order = c(1,0,0), ar = 0.4), n = 200)
Wt = exp(AR1)
```

### 2.a.
```{r}
# Plot of AR(1) process
plot(Wt, ylab = "Wt = e^Yt", main = 'AR(1)', type = 'o')
```


```{r}
acf(Wt)
```

```{r}
BoxCox.ar(Wt)
```
**Answer**:   
As Lambda=0 ==> Log transformation

```{r}
w1 <- log(Wt)
plot(w1, main="Transformed data (Lambda=0)", type = "o")
```

```{r}
acf(w1, main = 'After Log transformation')
```
**Answer**:   
It seems to be more stationary after the log transformation.
